---
title: "Evaluating multi-season occupancy models with autocorrelation fitted to heterogeneous datasets"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

André L Luza $^1$, Didier Alard $^{1,2}$, Frédéric Barraquand $^3$

$^1$ UMR Biodiversité Gènes et Communautés, University of Bordeaux,
INRAE, Pessac, France\
$^2$ US Fauna, University of Bordeaux, Pessac, France\
$^3$ Institute of Mathematics of Bordeaux, University of Bordeaux,
CNRS, Bordeaux INP, Talence, France\

Correspondence to:\
André L Luza: andre-luis.luza@u-bordeaux.fr\
Frédéric Barraquand: frederic.barraquand@u-bordeaux.fr\

# -----------------------------------------------

## Brief description of the files and folders in this repository:

#### Root \
|- *SuppInfo-A-D.pdf*: Supporting Information A, B, C and D associated to the article (also published as Online Supporting Information)\
|- *SuppInfo-E.pdf*: Supporting Information E: Assessment of model misspeficication, analysis of $\psi_{it} \times p_it$, and evaluation of the study scenario 3-2 with NNGP=15 spatial neighbors.\
|- *SuppInfo-F.pdf*: Supporting Information E: analyses using the Random Walk Occupancy Model by Outwaithe et al. (2018)\
|- *occ-model-rwalk-RE-simulations.txt*: Model of Outwaithe et al. (2018) in BUGS language\
|- *SuppInfo-G.pdf*: Supporting Information F: analyses using data of the four remaining species: the meadow brown, the false ringlet, the marsh fritillary, and the small copper\
|- *SuppInfo-H.pdf*: Supporting Information G: names/IDs of data contributors, as appearing in FAUNA data sets\
\

#### |- *Data*: Spatial data\
|-------- SpatialData: systematic grid (grid cells 1x1 km, French geographic system) from Observatoire FAUNA. The $1 \times 1$km spatial grid was downloaded from: https://observatoire-fauna.fr/ressources/publications?typePublication%5B%5D=fauna&themesID%5B%5D=6\

\

#### |- *Processed_data*: Data obtained after processing raw data sets \
|-------- AltitudeHabitatStats.RData: Altitude data (from EUDEM) and habitat data (from CORINE land use data)\
|-------- SpeciesData.RData: Occupancy data used in SDMs (from Observatoire FAUNA data base)\
|-------- Water.RData: CORINE land cover data regarding water vs non-water sites\
\

#### |- *R*: R scripts\

<span style="color:blue">
|--- *SIMULATIONS*: R scripts used in simulations. They are numbered sequentially, from 1 to 9. Thus, they should be run respecting this sequence:\
</span>
|-------- SIM-1-study1-sc0-... .R: Scripts to reproduce original work of Doser & Stoudt 2024 ("D&S") using Bernoulli sampling design. "c1": generate simulated data; "c2": apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures. Two separate folders that will host simulation results (path: model_output/output_simulations) and figures are created here. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> sims_D&S".\
|-------- SIM-2-study1-sc1-... .R: Scripts to simulate data with higher spatial autocorrelation (lower values of $\phi$), under the original Bernoulli sampling design of D&S. "c1": generate simulated data; "c2": apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> smooths_sims_D&S".\
|-------- SIM-3-a-study2-Sim-Settings.R: Our simulation settings with initial number of secondary sampling occasions $J=10$ (equivalent to sampling months in our butterfly data). These settings are used in Study 2. The code also create folders to store simulation results, figures, and the theme for ggplot (as per D&S).\
|-------- SIM-3-study2-sc0-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0). "c1": creates the Poisson sampling design - array $G_{itj}$; "c2": generate simulated data, apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_zero".\
|-------- SIM-4-study2-sc1-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy covariate. "c1": generate simulated data, apply the Poisson sampling design already generated in the previous scenario, and fit the multi-season occupancy model; "c2": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_one".\
|-------- SIM-5-study2-sc2-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate (total overlap of covariates). "c1": generate simulated data, apply the Poisson sampling design already generated in the previous scenario, and fit the multi-season occupancy model; "c2": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_two".\
|-------- SIM-6-study2-sc3-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate, and $v_{itj}$ (random covariate) in the detection model (partial overlap of covariates). "c1": generate simulated data, apply the Poisson sampling design already generated in the previous scenario, and fit the multi-season occupancy model; "c2": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_three".\
|-------- SIM-7-study3-sc1-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate, and $v_{itj}$ (random covariate) in the detection model (partial overlap of covariates). Design including the effect of phenology + observer preferences on occupancy data. "c1": creates the Poisson sampling design - array $G_{itj}$ - with an effect of phenology + observer preferences; "c2": generate simulated data, apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_phenology".\
|-------- SIM-8-study3-sc2-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate, and $v_{itj}$ (random covariate) in the detection model (partial overlap of covariates). Design including the effect of phenology + observer preferences + observation spot on occupancy data. "c1": creates the Poisson sampling design - array $G_{itj}$ - with an effect of phenology + observer preferences + observation spot in mid latitudes; "c2": generate simulated data, apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_phenology_spot".\
|-------- SIM-9-study3-sc2-...NNGP-15.R: Scripts used to analyze simulated data, as per study scenario 3-2, using NNGP=15 spatial neighbors in the Gaussian Process approximation. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_phenology_spot-NNGP-15-review".\
|-------- SIM-10-study3-sc3-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate, and $v_{itj}$ (random covariate) in the detection model (partial overlap of covariates). Design including the effect of phenology + observer preferences + observation spot of I*50% sites on occupancy data. "c1": creates the Poisson sampling design - array $G_{itj}$ - with an effect of phenology + observer preferences + observation spot in mid latitudes; "c2": generate simulated data, apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_phenology_spot_review".\
|-------- SIM-11-study2-sc1-... .R: Scripts to analyze data from study scenario 2-1 with the random walk model by Outhwaite et al. (2018) - https://doi.org/10.1016/j.ecolind.2018.05.010. "c1": build the model in BUGS language, generate simulated data, apply the Poisson sampling design and fit the multi-season occupancy model; "c2": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_two_sparta".\
|-------- SIM-12-study-simulation-comparison-Identifiability: Script used to interpret the output of simulations. This script will produce figures shown in the main text and SI\
|-------- SIM-12-simulation-psi-p-product: Evaluation of the product of $\psi_{it} \times p_{it}$ for each scenario.\

\

<span style="color:blue">
|--- *EMPIRICAL DATA ANALYSES*: R scripts used in the analyses of empirical data. They have the prefix "EMP" and are numbered sequentially, from 4 to 11. Code 1 to 3 involve the handling of sensitive information about threatened species and are not shown in the repository; code 4 from to 11 handle processed data used in our analyzes. Code should be run respecting this sequence:\
</span>
|-------- EMP-4-plot-occupancy-data.R: script used to plot butterfly data (Fig. 1) with occupancy already available in the folder "Processed_data" (raw data contain sensitive data. Access to raw data must be requested from Observatoire FAUNA, https://observatoire-fauna.fr/). The folder that will store figures will be created here.\
|-------- EMP-5-model-fit-buffer-L-dispar-stPGOcc.R: script used to fit models to large copper *Lycaena dispar* data at Bordeaux + 10 km buffer scale.\
|-------- EMP-5-model-fit-buffer-L-dispar-stPGOcc.R: script used to fit models to common blue *Polyommatus icarus* data at Bordeaux + 10 km buffer scale.\
|-------- EMP-6-buffer-Predictions-Maps-L-dispar.R: script used to make predictions and map the estimated distribution of the large copper at buffer scale.\
|-------- EMP-6-buffer-Predictions-Maps-P-icarus.R: script used to make predictions and map the estimated distribution of the common blue at buffer scale.\
|-------- EMP-7-buffer-Tables-PPO-L-dispar.R: script used to calculate summary statistics of regression coefficients, including the Prior-Posterior overlap (Tables in Supporting Information), from the model fitted to large copper data at buffer scale.\
|-------- EMP-7-buffer-Tables-PPO-P-icarus.R: script used to calculate summary statistics of regression coefficients, including the Prior-Posterior overlap (Tables in Supporting Information), from the model fitted to common blue data at buffer scale.\
|-------- EMP-8-model-fit-NouvAquit-L-dispar-stPGOcc.R: script used to fit models to large copper *Lycaena dispar* data at Nouvelle Aquitaine scale.\
|-------- EMP-8-model-fit-NouvAquit-P-icarus-stPGOcc.R: script used to fit models to common blue *Polyommatus icarus* data at Nouvelle Aquitaine scale.\
|-------- EMP-9-NouvAquit-Predictions-Maps-L-dispar.R: script used to make predictions and map the estimated distribution of the large copper at the Nouvelle Aquitaine scale.\
|-------- EMP-9-NouvAquit-Predictions-Maps-P-icarus.R: script used to make predictions and map the estimated distribution of the common blue at the Nouvelle Aquitaine scale.\
|-------- EMP-10-NouvAquit-Tables-PPO-L-dispar.R: script used to calculate summary statistics for regression coefficients, including the Prior-Posterior overlap (Tables in Supporting Information), from the model fitted to large copper data at the Nouvelle Aquitaine scale.\
|-------- EMP-10-NouvAquit-Tables-PPO-P-icarus.R: script used to calculate summary statistics for regression coefficients, including the Prior-Posterior overlap (Tables in Supporting Information), from the model fitted to common blue data at the Nouvelle Aquitaine scale.\
|-------- EMP-11-buffer-model-fit-remaining-species.R: script used to fit models to data of *Coenonympha oedippus*, *Euphydryas aurinia*, *Lycaena phlaeas*, and *Maniola jurtina* at Bordeaux + 10 km buffer scale.\
|-------- EMP-11-buffer-Predictions-Maps-remaining-species.R: script used to make predictions and map distribution of *Coenonympha oedippus*, *Euphydryas aurinia*, *Lycaena phlaeas*, and *Maniola jurtina* at Bordeaux + 10 km buffer scale\
|-------- EMP-11-buffer-Tables-PPO-remaining-species.R: script used to calculate summary statistics for regression coefficients, including the Prior-Posterior overlap (Tables in Supporting Information F), from models fitted to  *Coenonympha oedippus*, *Euphydryas aurinia*, *Lycaena phlaeas*, and *Maniola jurtina* at Bordeaux + 10 km buffer scale.\
\

#### |- *model_output*: General simulation settings, sampling designs, and results of the analyses of simulated and empirical data sets. These folders and files (RData) will be created while running scripts:
|--------------- *empirical*: results of empirical data analysis (model output and predictions) using the $spOccupancy$ package; \
|--------------- *output_simulations*: simulation settings and results: \
|---------------------- *"Xp_itj.rda"*: observation covariate $v_{itj}$ (generated once) \
|---------------------- *"sim-settings.rda"*: simulation constants and MCMC settings \
|---------------------- *"sampling_design_Poisson"*: sampling design used in study 2 from scenario 0 to 3\
|---------------------- *"sampling_design_Poisson_phenology"*: sampling design used in study 3 - scenario 1 (only phenology + observer preferences).\
|---------------------- *"sampling_design_Poisson_phenology_spot"*: sampling design used in study 3 - scenario 2 (phenology + observer preferences + observation spot I*25% of the sites).\
|---------------------- *"sampling_design_Poisson_phenology_spot_st3_sc3"*: sampling design used in study 3 - scenario 2 (phenology + observer preferences + observation spot I*50% of the sites).\
|---------------------- *sims_D&S*: correct data and simulation results as per Doser & Stoudt (2024) - **study 1-scenario 0** \
|---------------------- *smooth_sims_D&S*: correct data and simulation results for analysis of the scenario of high spatial autocorrelation - low spatial decay parameter $\phi$ - **study 1-scenario 1** \
|---------------------- *scenario_zero*: simulation results for study 2 - scenario 0 - **study 2-scenario 0** \
|---------------------- *scenario_one*: simulation results for study 2 - scenario 1 - **study 2-scenario 1** \
|---------------------- *scenario_two*: simulation results for study 2 - scenario 2 - **study 2-scenario 2** \
|---------------------- *scenario_two_sparta*: correct data, true $\psi_{it}$ from **study 2-scenario 1**, and simulation results for study 2 - scenario 1 using the SPARTA approach (uncorrelated site random effects) (Outwhaite et al. 2018) - **study 2-scenario 1** \
|---------------------- *scenario_three*: simulation results for study 2 - scenario 3 - **study 2-scenario 3** \
|---------------------- *scenario_phenology*: simulation results for the study 3 - scenario 1 incorporating phenology + observer preferences - **study 3 - scenario 1** \
|---------------------- *scenario_phenology_spot*: correct data and simulation results for study 3 - scenario 2 incorporating phenology + observer preferences + observation spot - **study 3 - scenario 2** \
|---------------------- *scenario_phenology_spot-NNGP-15-review*: analysis of simulated data for study 3 - scenario 2 with NNGP=15 spatial neighbors.\
|---------------------- *scenario_phenology_spot_review*: correct data and simulation results for study 3 - scenario 3 incorporating phenology + observer preferences + observation spot of 50% of the sites - **study 3 - scenario 3** \

\

#### |- *R-mis-specification*: Code used in the assessment of model misspecification (three scenarios analyzed):
|-------- SIM-1-study1-sc0-... .R: Scripts to reproduce original work of Doser & Stoudt 2024 ("D&S") using Bernoulli sampling design. "c1": generate simulated data with mis.spec.type = "probit"; "c2": apply the sampling design and fit the multi-season occupancy model (with logit link function); "c3": interpretation of results and figures. Two separate folders that will host simulation results (path: model_output/output_simulations) and figures are created here. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> sims_D&S".\
|-------- SIM-3-a-study2-Sim-Settings.R: Our simulation settings with initial number of secondary sampling occasions $J=10$ (equivalent to sampling months in our butterfly data). These settings are used in Study 2. The code also create folders to store simulation results, figures, and the theme for ggplot (as per D&S).\
|-------- SIM-5-study2-sc2-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate (total overlap of covariates). "c1": generate simulated data with pnorm link for $\psi_{it}$ and $p_{itj}$, apply the Poisson sampling design already generated in the previous scenario, and fit the multi-season occupancy model (with logit link function); "c2": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_two".\
|-------- SIM-8-study3-sc2-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate, and $v_{itj}$ (random covariate) in the detection model (partial overlap of covariates). Design including the effect of phenology + observer preferences + observation spot on occupancy data. "c1": creates the Poisson sampling design - array $G_{itj}$ - with an effect of phenology + observer preferences + observation spot in mid latitudes; "c2": generate simulated data with pnorm link for $\psi_{it}$ and $p_{itj}$, apply the sampling design and fit the multi-season occupancy model (with logit link function); "c3": interpretation of results and figures. Results and figures of this study-scenario are stored in one specific folder in the path: "model_output --> output_simulations --> scenario_phenology_spot".\


#### |- *model_output_mis-specification*: Results of the assessment of model misspecification (three scenarios analyzed):
|--------------- *figures*: Scatter plots showing the relationship between true and estimated $\psi_{it}$ under a misspecified model; \
|--------------- *output_simulations*: simulation settings and misspecification results for the following studies - scenarios: \
|---------------------- *"sim-settings.rda"*: simulation constants and MCMC settings \
|---------------------- *sims_D&S*: correct data and simulation results as per Doser & Stoudt (2024) - **study 1-scenario 0** \
|---------------------- *scenario_two*: simulation results for study 2 - scenario 2 - **study 2-scenario 2** \
|---------------------- *scenario_phenology_spot*: correct data and simulation results for study 3 - scenario 2 incorporating phenology +  observer preferences + observation spot of I*25% sites.\

##### This article was produced using the following software and associated packages:

\
```{r pressure, echo=F,warning=F,message=F}

source("R/packages.R")
sessionInfo()


```

\

#### End of the description                      




