---
title: "Evaluating multi-season occupancy models with autocorrelation fitted to heterogeneous datasets"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

André L Luza $^1$, Didier Alard $^{1,2}$, Frederic Barraquand $^3$

$^1$ UMR Biodiversité Gènes et Communautés, University of Bordeaux,
INRAE, Pessac, France\
$^2$ Observatoire FAUNA, University of Bordeaux, Pessac, France\
$^3$ Institute of Mathematics of Bordeaux, University of Bordeaux,
CNRS, Bordeaux INP, Talence, France

# -----------------------------------------------

## Brief description of the files and folders in this repository:

#### Root \
|- *SuppInfo-A-D.pdf*: Supporting Information A, B, C and D associated to the article (also published as Online Supportiong Information)\
|- *SuppInfo-E.pdf*: Supporting Information E: analyses using the Random Walk Occupancy Model by Outwaithe et al. (2018)\
|- *occ-model-rwalk-RE-simulations.txt*: Model of Outwaithe et al. (2018) in BUGS language\
|- *SuppInfo-F.pdf*: Supporting Information F: analyses using data of the four remaining species: the meadow brown, the false ringlet, the marsh fritillary, and the small copper\
|- *SuppInfo-G.pdf*: Supporting Information G: names/IDs of data contributors, as appearing in FAUNA data sets\
\

#### |- *Data*: Spatial data\
|-------- SpatialData: systematic grid (grid cells 1x1 km, French geographic system) from Observatoire FAUNA. The $1 \times 1$km spatial grid was downloaded from: https://observatoire-fauna.fr/ressources/publications?typePublication%5B%5D=fauna&themesID%5B%5D=6\

\

#### |- *Processed_data*: Data obtained after processing raw data sets \
|-------- AltitudeHabitatStats.RData: Altitude data (from EUDEM) and habitat data (from CORINE land use data)\
|-------- SpeciesData.RData: Occupancy data used in SDMs (from Observatoire FAUNA data base)\
|-------- Water.RData: CORINE land cover data regarding water vs non-water sites\
\

#### |- *R*: R scripts\

<span style="color:blue">
|--- *Simulations*: R scripts used in simulations. They are numbered sequentially, from 1 to 9. Thus, they should be run respecting this sequence:\
</span>
|-------- 1-study1-sc0-... .R: Scripts to reproduce original work of Doser & Stoudt 2024 = D&S, using Bernoulli sampling design. "c1": generate simulated data; "c2": apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures. Two separate folders that will host simulation results (path: model_output/output_simulations) and figures are created here. Results and figures of each study-scenario will be stored in one specific folder.\
|-------- 2-study1-sc1-... .R: Scripts to simulate data with higher spatial autocorrelation (lower values of $\phi$), under the original Bernoulli sampling design of D&S. "c1": generate simulated data; "c2": apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures.\
|-------- 3-a-study2-Sim-Settings.R: Our simulation settings with initial number of secondary sampling occasions $J=10$ (equivalent to sampling months in our butterfly data). These settings are used in Study 2. The code also create folders to store simulation results, figures, and the theme for ggplot.\
|-------- 3-study2-sc0-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0). "c1": creates the Poisson sampling design - array $G_{itj}$; "c2": generate simulated data, apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures\
|-------- 4-study2-sc1-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy covariate. "c1": generate simulated data, apply the Poisson sampling design already generated in the previous scenario, and fit the multi-season occupancy model; "c2": interpretation of results and figures\
|-------- 5-study2-sc2-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate (total overlap of covariates). "c1": generate simulated data, apply the Poisson sampling design already generated in the previous scenario, and fit the multi-season occupancy model; "c2": interpretation of results and figures\
|-------- 6-study2-sc3-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate, and $v_{itj}$ (random covariate) in the detection model (partial overlap of covariates). "c1": generate simulated data, apply the Poisson sampling design already generated in the previous scenario, and fit the multi-season occupancy model; "c2": interpretation of results and figures\
|-------- 7-study3-sc1-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate, and $v_{itj}$ (random covariate) in the detection model (partial overlap of covariates). Design including the effect of phenology + observer preferences on occupancy data. "c1": creates the Poisson sampling design - array $G_{itj}$ - with an effect of phenology + observer preferences; "c2": generate simulated data, apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures\
|-------- 8-study3-sc2-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and latitude $L_i$ as occupancy and detection covariate, and $v_{itj}$ (random covariate) in the detection model (partial overlap of covariates). Design including the effect of phenology + observer preferences + observation spot on occupancy data. "c1": creates the Poisson sampling design - array $G_{itj}$ - with an effect of phenology + observer preferences + observation spot in mid latitudes; "c2": generate simulated data, apply the sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures\
|-------- 9-study2-sc1-... .R: Scripts to simulations with our Poisson sampling design ($\phi$ as in study 1-sc0) and combination of covariates as study2-sc1 - latitude $L_i$ as occupancy covariate, $v_{itj}$ as detection covariate. Here the data was analyzed using the random walk model by Outhwaite et al. (2018) - https://doi.org/10.1016/j.ecolind.2018.05.010. "c1": build the model in BUGS language, generate simulated data, apply the Poisson sampling design and fit the multi-season occupancy model; "c3": interpretation of results and figures\
|-------- 10-study-simulation-comparison-Identifiability: Script used to interpret the output of simulations. This script will produce figures shown in the main text and SI\
\

<span style="color:blue">
|--- *Empirical analyses*: R scripts used in the analyses of empirical data. They are numbered sequentially, from 11 to 20. Thus, they should be run respecting this sequence:\
</span>
|-------- 11-study-empirical-plot-occupancy-data.R: script used to plot butterfly data (Fig. 1) with occupancy already available in the folder "Processed_data" (raw data contain sensitive data. Access to raw data must be requested from Observatoire FAUNA, https://observatoire-fauna.fr/). The folder that will store figures will be created here.\
|-------- 12-study-empirical-buffer-DataAnalysis-... .R: script used to analyze data at Bordeaux + 10 km buffer scale. One script per species -  *P. icarus* and *L. dispar*.\
|-------- 13-study-empirical-buffer-Predictions-Maps-... .R: script used to make predictions and map estimated distribution at buffer scale. One script per species\
|-------- 14-study-empirical-buffer-Tables-PPO-... .R: script used to calculate summary statistics for regression coefficients, including the Prior-Posterior overlap (Tables in Supporting Information). One script per species\
|-------- 15-study-empirical-buffer-DataAnalysis-... .R: script used to analyze data at Nouvelle Aquitaine scale. One script per species\
|-------- 16-study-empirical-buffer-Predictions-Maps-... .R: script used to make predictions and map estimated distribution at Nouvelle Aquitaine scale. One script per species\
|-------- 17-study-empirical-buffer-Tables-PPO--... .R: script used to calculate summary statistics for regression coefficients, including the Prior-Posterior overlap (Tables in Supporting Information) at Nouvelle Aquitaine scale. One script per species\
|-------- 18-study-empirical-buffer-DataAnalysis-remaining-spp-stPGocc.R: script used to analyze data of the four remaining species - Bordeaux + 10 km buffer scale\
|-------- 19-study-empirical-buffer-Predictions-Maps-remaining-species.R: script used to make predictions and map distribution of the remaining species - Bordeaux + 10 km buffer scale\
|-------- 20-study-empirical-buffer-Tables-PPO-remaining-species.R: script used to calculate summary statistics for regression coefficients, including the Prior-Posterior overlap (Tables in Supporting Information F). Results shown only at Bordeaux + 10 km buffer scale\
\

#### |- *model_output*: results of the analyses of simulated and empirical data sets - these folders and files (RData) will be created while running scripts:
|--------------- *empirical*: results of empirical data analysis (model output and predictions) using the $spOccupancy$ package; \
|--------------- *output_simulations*: simulation settings and results: \
|---------------------- *"Xp_itj.rda"*: observation covariate $v_{itj}$ (generated once) \
|---------------------- *"sim-settings.rda"*: simulation constants and MCMC settings \
|---------------------- *"sampling_design_Poisson"*: sampling design used in study 2 from scenario 0 to 3\
|---------------------- *"sampling_design_Poisson_phenology"*: sampling design used in study 3 - scenario 1 (only phenology + observer preferences)\
|---------------------- *"sampling_design_Poisson_phenology_spot"*: sampling design used in study 3 - scenario 2 (phenology + observer preferences + observation spot) \
|---------------------- *sims_D&S*: correct data and simulation results as per Doser & Stoudt (2024) - **study 1-scenario 0** \
|---------------------- *smooth_sims_D&S*: correct data and simulation results for analysis of the scenario of high spatial autocorrelation - low spatial decay parameter $\phi$ - **study 1-scenario 1** \
|---------------------- *scenario_zero*: simulation results for study 2 - scenario 0 - **study 2-scenario 0** \
|---------------------- *scenario_one*: simulation results for study 2 - scenario 1 - **study 2-scenario 1** \
|---------------------- *scenario_two*: simulation results for study 2 - scenario 2 - **study 2-scenario 2** \
|---------------------- *scenario_two_sparta*: correct data, true $\psi_{it}$ from **study 2-scenario 1**, and simulation results for study 2 - scenario 1 using the SPARTA approach (uncorrelated site random effects) (Outwhaite et al. 2018) - **study 2-scenario 1** \
|---------------------- *scenario_three*: simulation results for study 2 - scenario 3 - **study 2-scenario 3** \
|---------------------- *scenario_phenology*: simulation results for the study 3 - scenario 1 incorporating phenology + observer preferences - **study 3 - scenario 1** \
|---------------------- *scenario_phenology_spot*: correct data and simulation results for study 3 - scenario 2 incorporating phenology + observer preferences + observation spot - **study 3 - scenario 2** \
\

##### This article was produced using the following software and associated packages:

\
```{r pressure, echo=F,warning=F,message=F}

source("R/packages.R")
sessionInfo()


```

\

#### End of the description                      




